---
title: "Demographics (patient information)"
---

The cleaning script used here is at: [GitHub: mover, 01_patient_information.R](https://github.com/SurgicalInformatics/mover/blob/main/01_patient_information.R)

```{r}
library(tidyverse)
library(finalfit)
source("functions.R")
theme_set(theme_bw())

demographics = read_csv("/home/common/mover_data/surginf_cleaned/patient_information_cleaned.csv")
demographics_orig = read_csv(paste0(Sys.getenv("epic_emr"), "patient_information.csv"), col_select = c("MRN", "LOG_ID"))

```

The patient information dataset includes `r n_distinct(demographics_orig$MRN)` patients who had `r n_distinct(demographics_orig$LOG_ID)` surgeries. The original file has `r nrow(demographics_orig)` rows, removing duplicate patient ID (`mrn`) and procedure ID (`log_id`) rows, leaves us with `r nrow(demographics)` patient-procedure observations.


Variable list: [MOVER: Patient Info Table](https://mover.ics.uci.edu/patient-info-table.html)


```{r}
demographics %>% 
  mutate(across(all_of(c("sex", "primary_anes_type_nm", "patient_class_group")), fct_infreq)) %>% 
  summary_factorlist(explanatory = c("age", "sex", "los", "primary_anes_type_nm", "asa_rating", "patient_class_group")) %>% 
  kable()
```


# Length of stay

```{r}
demographics %>% 
  summary_factorlist("los", "patient_class_group", cont = "median") %>% 
  kable()
```

```{r, fig.width=10, fig.height=10}
demographics %>% 
  drop_na(los) %>% 
  ggplot(aes(los, fill = patient_class_group)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(if_else(los>50, "LOS > 50 days", "LOS < 50 days")~patient_class_group, scales = "free", ncol = 1) +
  scale_x_continuous() +
  scale_fill_viridis_d() +
  theme(legend.position =  c(0.9, 0.95))
```


# Procedures


```{r}
demographics %>% 
  count(primary_procedure_nm, sort = TRUE) %>% 
  # Changing from ALL CAPS to lowercase
  mutate(primary_procedure_nm = str_to_sentence(primary_procedure_nm)) %>% 
  mydt(pagelength = 50)
```

